---
title: "Problem Set 7 -- MATH392"
author: "Theodore Dounias"
date: "3/25/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(resampledata)
library(knitr)
library(kableExtra)
library(broom)

data <- corrExerciseB
```

####9.4
     
If I consider -5Y to be an rv of its own, I can apply the following identity:
$$Var[2X - 5Y] = Var[2X] + Var[-5Y] + 2Cov[2X, -5Y]$$   
     
To find the covariance, I will use the following:
   
$$Cov[2X, -5Y] = E[-10XY] - E[2X]E[-5Y] = -10(E[XY] - E[X]E[Y])= -10Cov[X, Y] = -20$$
    
Plugging in to the first equation I have:
$$Var[2X-5Y] = 12 + 150 - 20 = 142$$
    
    
####9.7
     
A.
```{r}
cor(data$X, data$Y)
```
     
B.    
```{r, warning=FALSE}
data <- data %>%
  group_by(Z) %>%
  summarize(xbar = mean(X), ybar = mean(Y))

data %>%
  kable() %>%
  kable_styling()
```
     
C.    
```{r, fig.height=2, fig.width = 3, fig.align='center'}
ggplot(data, aes(x = xbar, y = ybar)) +
  geom_text(aes(label = Z)) +
  geom_line(alpha = .5, col = "skyblue")
```
    
####9.9
     
I will use the fact that $\beta_1 = \bar y-\beta_2 \bar x$ and $\hat y = \beta_1 + \beta_2 x$:
    
$$\sum_{i = 1}^{n} (\hat y - y_i) = \sum_{i = 1}^{n} (y_i - (\beta_1 + \beta_2 x_i)) = -\beta_1n +\sum_{i = 1}^{n} (y_i -\beta_2 x_i) = -\beta_1 n + n\sum_{i = 1}^{n} (\frac{y_i}{n}-\beta_2 \frac{x_i}{n})$$   
     
By using the aforementioned equations I have:
    
$$\sum_{i = 1}^{n} (\hat y - y_i) = -n(\bar y-\beta_2 \bar x) + n(\bar y-\beta_2 \bar x) = 0$$
     
####9.14
     
A.     
```{r, warning= FALSE, fig.height=2, fig.width = 3, fig.align='center'}
data <- Illiteracy
   

ggplot(data, aes(y = Illit, x = Births)) +
  geom_point() +
  geom_smooth(alpha = .02)
```
    
The relationship appears to be linear; higher birth rates and higher illiteracy increase simultaneously.
     
B.    
```{r}
lm1 <- lm(data = data, Illit ~ Births)

tidy(lm1)

glance(lm1)[1]
```
    
The slope coefficient is statisticaly significant, positive, and equal to 10.83; this means that for an increase of one birth, there is approximatelly a 10.83 increase in illiteracy rates. The residual sum of squares is about .6, which signifies the error rate that the output line has in predicting the data.
    
C.   
```{r, warning= FALSE, fig.height=2, fig.width = 3, fig.align='center'}

plot(lm1)
```
     
If we consider these diagnostic plots, it is obvious that the residuals present an increase in variance as the values of illiteracy themselves increase. This might suggest that using a log value for the response variable might be appropriate, but the linear model is a fine approximation.
     
D.  Not necessarily. To consider policy solutions we would have to run a wider array of models that include other variables. What we are seeing in this simple model might be an example of collinearity, a relationship that exists due to an underlying cause. In that case, we should try to find the true root of the issue.
