---
title: "Supplement--Midterm II"
author: "Theodore Dounias"
date: "3/8/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I will try to reach a parametric equasion that includes b, a, and n. I will then solve for each value of n that are asked for. First, from the cdf for a N(100, 15/sqrt(n)) distribution I have:
$$P(\bar X < q) = 1-a \Rightarrow P(\frac{\bar X - 100}{15/ \sqrt n}<\frac{q - 100}{15/ \sqrt n})=1-a\Rightarrow F_Z(\frac{q - 100}{15/ \sqrt n}) = 1-a \Rightarrow q = \frac{15}{\sqrt n}F_z^{-1}(1 - a) +100$$
    
Where $F_Z^{-1}$ is the inverse cdf of the standard normal. Assuming now that our initial X derives from a N(105, 15/sqrt(n)) as we did in class, we have a second equation:
    
$$F(q)=b\Rightarrow F_Z(\frac{q - 105}{15/ \sqrt n})=b \Rightarrow q = \frac{15}{\sqrt n}F_z^{-1}(b) + 105$$
    
If we equate and solve for n and alpha = .05 we have:
$$n = 9(F_z^{-1}(1-a) - F_z^{-1}(b))^2 \Rightarrow n = 9(1.64 - F_z^{-1}(b))^2$$
   
I can also generalize this equasion to:
$$n = \frac{\sigma^2}{(difference~in~means)^2}(F_z^{-1}(1-a) - F_z^{-1}(b))^2$$
   
Using r we have:
```{r}
n_power <- function(a, b){
  9*(qnorm(1-a) - qnorm(b))^2
}

ceiling(n_power(.05, .3))
ceiling(n_power(.05, .2))
ceiling(n_power(.05, .05))
```

